---
title: "Clean Data in R"
output: html_notebook
author:  Jeff Gross
---

#Packages
```{r}
install.packages(dplyr)
install.packages("tidyr")
install.packages("lubridate")

#get a def of a function
getAnywhere(draw_roc_lines)

library(dplyr)
library(tidyr)
library(lubridate)
```


#Getting a feel for your data

We've loaded a dataset called bmi into your workspace. The data, which give the (age standardized) mean body mass index (BMI) among males in each country for the years 1980-2008, come from the School of Public Health, Imperial College London.

Check the class of bmi
Find the dimensions of bmi
Print the bmi column names
```{r}
bmi <- read_csv("~/R Scripts/bmi.csv")

# Check the class of bmi
class(bmi)

# Check the dimensions of bmi
dim(bmi)

# View the column names of bmi
names(bmi)
```
#Viewing the structure of your data

View the structure of bmi using the traditional method
Load the dplyr package
View the structure of bmi using dplyr
Look at a summary() of bmi
```{r}
# Check the structure of bmi
str(bmi)

# Load dplyr
library(dplyr)

# Check the structure of bmi, the dplyr way
glimpse(bmi)

# View a summary of bmi
summary(bmi)
```
#Looking at your data

Print the full dataset to the console (you don't need print() to do this)
View the first 6 rows of bmi
View the first 15 rows of bmi
View the last 6 rows of bmi
View the last 10 rows of bmi
```{r}
# Print bmi to the console
#print(bmi)

# View the first 6 rows
head(bmi)

# View the first 15 rows
head(bmi, n=15)

# View the last 6 rows
tail(bmi)

# View the last 10 rows
tail(bmi, n=10)
```
#Visualizing your data

For the bmi dataset:

Use hist() to look at the distribution of average BMI across all countries in 2008
Use plot() to see how each country's average BMI in 1980 (x-axis) compared with its BMI in 2008 (y-axis)
```{r}
# Histogram of BMIs from 2008
hist(bmi$Y2008)

# Scatter plot comparing BMIs from 1980 to those from 2008
plot(bmi$Y1980,bmi$Y2008)
```
#Gathering columns into key-value pairs

gather(wide_df, my_key, my_val, -col)

Apply the gather() function to bmi, saving the result to bmi_long. This will create two new columns:
year, containing as values what are currently column headers
bmi_val, the actual BMI values
View the first 20 rows of bmi_long
```{r}
# Apply gather() to bmi and save the result as bmi_long
bmi_long <- gather(bmi, bmi_value, year, -Country)

# View the first 20 rows of the result
head(bmi_long, n=20)
```
#Spreading key-value pairs into columns

spread(long_df, my_key, my_val)

Use spread() to reverse the operation that you performed in the last exercise with gather(). In other words, make bmi_long wide again, saving the result to bmi_wide
View the head of bmi_wide
```{r}
# Apply spread() to bmi_long
bmi_wide <- spread(bmi_long, year, bmi_val)

# View the head of bmi_wide
head(bmi_wide)
```
#Separating columns

separate(treatments, year_mo, c("year", "month"))

```{r}
bmi_cc <- read_csv("~/R Scripts/bmi_cc.csv")

# Apply separate() to bmi_cc
bmi_cc_clean <- separate(bmi_cc, col = Country_ISO, into = c("Country", "ISO"), sep = "/")

# Print the head of the result
head(bmi_cc_clean)
```
#Uniting columns

unite(treatments, year_mo, year, month)

In the last exercise, you separated the Country_ISO column of the bmi_cc dataset into two columns (Country and ISO) and saved the result to bmi_cc_clean. Now you're going to put the columns back together!

Apply the unite() function to bmi_cc_clean
Reunite the Country and ISO columns into a single column called Country_ISO
Separate each country name and code with a dash (-)
Save the result as bmi_cc
View the head of the result
```{r}
# Apply unite() to bmi_cc_clean
bmi_cc <- unite(bmi_cc_clean, Country_ISO, Country, ISO, sep = "-")

# View the head of the result
head(bmi_cc)
```
#Column headers are values, not variable names

View the head of census.
Gather the month columns, creating two new columns (month and amount), saving the result to census2.
Run the code given to arrange() the rows of census2 by the YEAR column.
View the first 20 rows of the result.
Show Answer (-70xp)
Hint
Call gather() with four arguments:
The name of the data object
The name of the new column whose values will be what are now the column headers (i.e. the key argument)
The name of the new column whose values will be the month measurements (i.e. the amount argument)
The columns to gather (or, in this case, the column to exclude from gathering. Which column in census do we not want to gather()?)
```{r}
## tidyr and dplyr are already loaded for you

# View the head of census
head(census)

# Gather the month columns
census2 <- gather(census, month, amount, -YEAR)

# Arrange rows by YEAR using dplyr's arrange
census2 <- arrange(census2, YEAR)

# View first 20 rows of census2
head(census2, n=20)
```
#Variables are stored in both rows and columns

View the first 50 rows of census_long
Decide which column of census_long would be best to spread, and which column of census_long would be best to display in the newly spread out columns. Use the spread() function accordingly and save the result to census_long2
View the first 20 rows of census_long2
```{r}
pets <- read_csv("~/R Scripts/pets.csv")
census_long <- read_csv("~/R Scripts/census_long.csv")

head(pets)
spread(pets, type, num)

## tidyr is already loaded for you

# View first 50 rows of census_long
head(census_long, n=50)

# Spread the type column
census_long2 <- spread(census_long, type, amount)

# View first 20 rows of census_long2
head(census_long2, n=50)
```
#Multiple values are stored in one column

The separate() function comes in handy in these situations. To practice using it, we have created a slight modification of last exercise's result. Keep in mind that the into argument, which specifies the names of the 2 new columns being formed, must be given as a character vector (e.g. c("column1", "column2")).

View the head of census_long3
Use tidyr's separate() to split the yr_month column into two separate variables: year and month, saving the result to census_long4
View the first 6 rows of the result
```{r}
census_long3 <- read_csv("~/R Scripts/census_long3.csv")

## tidyr is already loaded for you

# View the head of census_long3
head(census_long3)

# Separate the yr_month column into two
census_long4 <- separate(census_long3,yr_month, c("year", "month"))

# View the first 6 rows of the result
head(census_long4)
```
#Types of variables in R

Change the object within each call of the class() function to make it evaluate to the following (in order):

character
numeric
integer
factor
logical
Add or remove quotes, add an L to numerics to make them integers and use the factor() function when appropriate to accomplish this.
```{r}
# Make this evaluate to character
class("true")

# Make this evaluate to numeric
class(8484.00)

# Make this evaluate to integer
class(99L)

# Make this evaluate to factor
class(factor("factor"))

# Make this evaluate to logical
class(FALSE)
```
#Common type conversions

There are a few less intuitive results. For example, under the hood, the logical values TRUE and FALSE are coded as 1 and 0, respectively. Therefore, as.logical(1) returns TRUE and as.numeric(TRUE) returns 1.

We've loaded a dataset called students into your workspace. These data provide information on 395 students including their grades in three classes (in the Grades column, separated by /).

Use str() to preview students and see the class of each variable
Coerce the following columns:
Grades to character
Medu to factor (categorical variable representing mother's education level)
Fedu to factor (categorical variable representing father's education level)
Use str() again to see the changes to students
```{r}
students <- read_csv("~/R Scripts/students.csv")

# Preview students with str()
str(students)

# Coerce Grades to character
students$Grades <- as.character(students$Grades)

# Coerce Medu to factor
students$Medu <- as.factor(students$Medu)

# Coerce Fedu to factor
students$Fedu <- as.factor(students$Fedu)
    
# Look at students once more with str()
str(students)
```
#Working with dates

We have loaded a dataset called students2 into your workspace. students2 is similar to students, except now instead of an age for each student, we have a (hypothetical) date of birth in the dob column. There's another new column called nurse_visit, which gives a timestamp for each student's most recent visit to the school nurse.

As you saw in the video, these functions combine the letters y, m, d, h, m, s, which stand for year, month, day, hour, minute, and second, respectively. The order of the letters in the function should match the order of the date/time you are attempting to read in, although not all combinations are valid. Notice that the functions are "smart" in that they are capable of parsing multiple formats.

Preview students2 with str(). Notice that dob and nurse_visit are both stored as character
Load the lubridate package
Print "17 Sep 2015" as a date
Print "July 15, 2012 12:56" as a date and time (note there are hours and minutes, but no seconds!)
Coerce dob to a date (with no time)
Coerce nurse_visit to a date and time
Use str() to see the changes to students2
```{r}
students2 <- read_csv("~/R Scripts/students2.csv")

# Preview students2 with str()
str(students2)

# Load the lubridate package
library(lubridate)

# Parse as date
dmy("17 Sep 2015")

# Parse as date and time (with no seconds!)
mdy_hm("July 15, 2012 12:56")

# Coerce dob to a date (with no time)
students2$dob <- ydm(students2$dob)

# Coerce nurse_visit to a date and time
students2$nurse_visit <- ymd_hms(students2$nurse_visit)
    
# Look at students2 once more with str()
str(students2)
```
#Trimming and padding strings

> str_trim("  this is a test     ")
[1] "this is a test"

> str_pad("24493", width = 7, side = "left", pad = "0")
[1] "0024493"

Load the stringr package
Trim all leading and trailing white space from the first set of strings
Pad the second set of strings with leading zeros such that all are 9 characters in length
```{r}
# Load the stringr package
library(stringr)

# Trim all leading and trailing whitespace
str_trim(c("   Filip ", "Nick  ", " Jonathan"))

# Pad these strings with leading zeros
str_pad(c("23485W", "8823453Q", "994Z"),width = 9, side = "left", pad= "0")
```
Examples like this are certainly handy in R. For example, the str_pad() function is useful when importing a dataset with US zip codes. Occasionally R will drop the leading 0 in a zipcode, thinking it's numeric. Now that you know how to coerce variable types and pad strings, this won't set you back.

#Upper and lower case

There's a vector of state abbreviations called states in your workspace, but there's a problem...it's all lowercase. It's more common for state abbreviations to be all uppercase.

Print states to the console
Make states all uppercase and save the result to states_upper
Make states_upper all lowercase again, but don't save the result
```{r}
states <- c("al",	"ak",	"az",	"ar",	"ca",	"co",	"ct",	"de",	"fl",	"ga",	"hi",	"id",	"il",	"in",	"ia",
"ks",	"ky",	"la",	"me",	"md",	"ma",	"mi",	"mn",	"ms",	"mo",	"mt",	"ne",	"nv",	"nh",	"nj",
"nm",	"ny",	"nc",	"nd",	"oh",	"ok",	"or",	"pa",	"ri",	"sc",	"sd",	"tn",	"tx",	"ut",	"vt",
"va",	"wa",	"wv",	"wi",	"wy")

# Print state abbreviations
states

# Make states all uppercase and save result to states_upper
states_upper <- toupper(states)

# Make states_upper all lowercase again
tolower(states_upper)
```
#Finding and replacing strings

The stringr package provides two functions that are very useful for finding and/or replacing strings: str_detect() and str_replace().

Like all functions in stringr, the first argument of each is the string of interest. The second argument of each is the pattern of interest. In the case of str_detect(), this is the pattern we are searching for. In the case of str_replace(), this is the pattern we want to replace. Finally, str_replace() has a third argument, which is the string to replace with.

Instructions
The students2 dataset from earlier in the chapter has been loaded for you again.

Look at the head() of students2 to remind yourself of how it looks
Detect all dates of birth (dob) in 1997 using str_detect(). This should return a vector of TRUE and FALSE values.
Replace all instances of "F" with "Female" in students2$sex
Replace all instances of "M" with "Male" in students2$sex
View the head() of students2 to see the result of these replacements
```{r}
## stringr has been loaded for you

# Look at the head of students2
head(students2)

# Detect all dates of birth (dob) in 1997
str_detect(students2$dob, "1997")

# In the sex column, replace "F" with "Female"...
students2$sex <- str_replace(students2$sex, "F", "Female")

# ...And "M" with "Male"
students2$sex <- str_replace(students2$sex, "M", "Male")

# View the head of students2
head(students2)
```
#Finding missing values

If missing values are properly coded as NA, the is.na() function will help you find them. Otherwise, if your dataset is too big to just look at the whole thing, you may need to try searching for some of the usual suspects like "", "#N/A", etc. You can also use the summary() and table() functions to turn up unexpected values in your data.

Call is.na() on social_df to spot all NA values.
Wrap the above with the any() function to ask the question "Are there any NA values in my dataset?".
View a summary() of the dataset to see how missing values are broken out.
Use table to identify odd values of the status variable.
```{r}
social_df <- read_csv("~/R Scripts/social_df.csv")
social_df$name <- as.factor(social_df$name)
social_df$status <- as.factor(social_df$status)

# Call is.na() on the full social_df to spot all NAs
is.na(social_df)

# Use the any() function to ask whether there are any NAs in the data
any(is.na(social_df))

# View a summary() of the dataset
summary(social_df)

# Call table() on the status column
table(social_df$status)
```
#Dealing with missing values

Missing values can be a rather complex subject, but here we'll only look at the simple case where you are simply interested in normalizing and/or removing all missing values from your data. For more information on why this is not always the best strategy, search online for "missing not at random."

Instructions
Replace all empty strings (i.e. "") with NA in the status column of social_df.
Print the updated version of social_df to confirm your changes.
Use complete.cases() to return a vector containing TRUE and FALSE to see which rows have NO missing values.
Use na.omit() to remove all rows with one or more missing values (without saving the result).
```{r}
## The stringr package is preloaded

# Replace all empty strings in status with NA
social_df$status[social_df$status == ""] <- NA

# Print social_df to the console
social_df

# Use complete.cases() to see which rows have no missing values
complete.cases(social_df)

# Use na.omit() to remove all rows with any missing values
na.omit(social_df)
```
#Dealing with outliers and obvious errors

Instructions
Call summary() on the full students3 dataset to expose the concerning values of age and absences.
View a histogram (using hist()) of the age variable.
View a histogram of the absences variable.
View another histogram of absences, but force values of zero to be bucketed to the right of zero on the x-axis with right = FALSE (see ?hist for more info).
```{r}
students3 <- read_csv("~/R Scripts/students3.csv")

# Look at a summary() of students3
summary(students3)

# View a histogram of the age variable
hist(students3$age)

# View a histogram of the absences variable
hist(students3$absences)

# View a histogram of absences, but force zeros to be bucketed to the right of zero
hist(students3$absences, right=FALSE)
```
#Another look at strange values

If you're curious about the exact formula for determining what is "far", check out ?hist.)

In this situation, we are concerned about three things:

Since this dataset is about students and the only student above the age of 22 is 38 years old, we must wonder whether this is an error in the data or just an older student (perhaps returning to school after working for several years)
There are four values of -1 for the absences variable, which is either a mistake or an intentional coding meant to say, for example, "this value is missing"
There are several extreme values of absences in the positive direction, with a maximum value of 75 (which is over 18 times the median value of 4)
Instructions
View a boxplot() of the age variable from students3
View a boxplot() of the absences variable from students3
```{r}
# View a boxplot of age
boxplot(students3$age)

# View a boxplot of absences
boxplot(students3$absences)
```
#Get a feel for the data

Instructions
For the weather dataset, which is loaded in your workspace:

Check that it's a data.frame using the function class()
Look at the dimensions
View the column names
```{r}
weather <- read_csv("~/R Scripts/weather.csv")

# Verify that weather is a data.frame
class(weather)

# Check the dimensions
dim(weather)

# View the column names
names(weather)
```
#Summarize the data

Instructions
View the structure of weather using base R
Load the dplyr package
View the structure of weather, the dplyr way
View a summary() of weather
```{r}
# View the structure of the data
str(weather)

# Load dplyr package
library(dplyr)

# Look at the structure using dplyr's glimpse()
glimpse(weather)

# View a summary of the data
summary(weather)
```
#Take a closer look

Instructions
For the weather data:

View the first 6 rows
View the first 15 rows
View the last 6 rows
View the last 10 rows
```{r}
# View first 6 rows
head(weather)

# View first 15 rows
head(weather, n=15)

# View the last 6 rows
tail(weather)

# View the last 10 rows
tail(weather, n=10)
```
#Column names are values

Instructions
Load the tidyr package
Call gather() on the weather data to gather columns X1-X31. The two columns created as a result should be called day and value. Save the result as weather2
View the result with head()
```{r}
df <- read_csv("~/R Scripts/df.csv")
df
df2 <- gather(df, time, val, t1:t3)

# Load the tidyr package
library(tidyr)

# Gather the columns
weather2 <- gather(weather, day, value, X1:X31, na.rm = TRUE)

# View the head
head(weather2)
```
#Values are variable names

Our data suffer from a second common symptom of messy data: values are variable names. Specifically, values in the measure column should be variables (i.e. column names) in our dataset.

The spread() function from tidyr is designed to help with this. To remind you of how this function works, we've loaded another small dataset called df2 (which is the result of applying gather() to the original df from last exercise). Give the following a try before attempting the instructions below.
```{r}
df2
spread(df2, time, val)

## The tidyr package is already loaded

# First remove column of row names
weather2 <- weather2[, -1]

# Spread the data
weather3 <- spread(weather2, measure, value)

# View the head
head(weather3)
```
#Clean up dates

Instructions
Load the stringr and lubridate packages
Use stringr's str_replace() to remove the Xs from the day column of weather3
Create a new column called date. Use the unite() function from tidyr to paste together the year, month, and day columns in order, using - as a separator (see ?unite if you need help)
Coerce the date column using the appropriate function from lubridate
Use the code provided (select()) to reorder columns, saving the result to weather5
View the head of weather5
```{r}
## tidyr and dplyr are already loaded

# Load the stringr and lubridate packages
library(stringr)
library(lubridate)

# Remove X's from day column
weather3$day <- str_replace(weather3$day, "X","")

# Unite the year, month, and day columns
weather4 <- unite(weather3, date, year, month, day, sep = "-")

# Convert date column to proper date format using lubridates's ymd()
weather4$date <- ymd(weather4$date)

# Rearrange columns using dplyr's select()
weather5 <- select(weather4, date, Events, CloudCover:WindDirDegrees)

# View the head of weather5
head(weather5)
```
#A closer look at column types


```{r}
as.numeric(c(4, 6.44, "some string", 222))

# View the structure of weather5
str(weather5)

# Examine the first 20 rows of weather5. Are most of the characters numeric?
head(weather5, n=20)

# See what happens if we try to convert PrecipitationIn to numeric
as.numeric(weather5$PrecipitationIn)
```
If you run the code above in the console, you'll get a warning message saying that R introduced an NA in the process of coercing to numeric. This is because it doesn't know how to make a number out of a string ("some string"). Watch out for this in our weather data!

#Column type conversions

As you saw in the last exercise, "T" was used to denote a trace amount (i.e. too small to be accurately measured) of precipitation in the PrecipitationIn column. In order to coerce this column to numeric, you'll need to deal with this somehow. To keep things simple, we will just replace "T" with the number zero.

Instructions
Use str_replace() from stringr to make the proper replacements in the PrecipitationIn column of weather5
Run the call to mutate_each as-is to conveniently apply as.numeric() to all columns from CloudCover through WindDirDegrees (reading left to right in the data), saving the result to weather6
View the structure of weather6 to confirm the coercions were successful
```{r}
## The dplyr and stringr packages are already loaded
class(weather5$PrecipitationIn)

# Replace T with 0 (T = trace)
weather5$PrecipitationIn <- str_replace(weather5$PrecipitationIn, "T", "0")

# Convert characters to numerics
weather6 <- mutate_each(weather5, funs(as.numeric), CloudCover:WindDirDegrees)

# Look at result
str(weather6)
```
#Find missing values

Instructions
Use sum() and is.na() to count the number of NA values in weather6
Look at a summary() of weather6 to figure out how the missings are distributed among the different variables
Use which() to identify the indices (i.e. row numbers) where Max.Gust.SpeedMPH is NA and save the result to ind (for indices)
Use ind to look at the full rows of weather6 for which Max.Gust.SpeedMPH is missing
```{r}
# Count missing values
sum(is.na(weather6))

# Find missing values
summary(weather6)

# Find indices of NAs in Max.Gust.SpeedMPH
ind <- which(is.na(weather6$Max.Gust.SpeedMPH))

# Look at the full rows for records missing Max.Gust.SpeedMPH
weather6[ind, ]
```
#An obvious error

Instructions
View a summary() of weather6
Use which() to find the index of the erroneous element of weather6$Max.Humidity, saving the result to ind
Use ind to look at the full row of weather6 for that day
You discover an extra zero was accidentally added to this value. Correct it in the data
```{r}
# Review distributions for all variables
summary(weather6)

# Find row with Max.Humidity of 1000
ind <- which(weather6$Max.Humidity==1000)

# Look at the data for that day
weather6[ind, ]

# Change 1000 to 100
weather6$Max.Humidity[ind] <- 100
```
#Another obvious error

Instructions
Use summary() to look at the value of only the Mean.VisibilityMiles variable of weather6
Determine the element of the value that is clearly erroneous in this column, saving the result to ind
Use ind to look at the full row of weather6 for this day
Inspect the values of other variables for this day to determine the correct value of Mean.VisibilityMiles, then make the appropriate fix
```{r}
# Look at summary of Mean.VisibilityMiles
summary(weather6$Mean.VisibilityMiles)

# Get index of row with -1 value
ind <- which(weather6$Mean.VisibilityMiles==-1)

# Look at full row
weather6[ind,]

# Set Mean.VisibilityMiles to the appropriate value
weather6$Mean.VisibilityMiles[ind] <- 10
```
#Check other extreme values

Instructions
Check a summary() of weather6 one more time for extreme or unexpected values
View a histogram for MeanDew.PointF
Do the same for Min.TemperatureF
And once more for Mean.TemperatureF to compare distributions
```{r}
# Review summary of full data once more
summary(weather6)

# Look at histogram for MeanDew.PointF
hist(weather6$MeanDew.PointF)

# Look at histogram for Min.TemperatureF
hist(weather6$Min.TemperatureF)

# Compare to histogram for Mean.TemperatureF
hist(weather6$Mean.TemperatureF)
```
#Finishing touches

Instructions
We've created a vector of column names in your workspace called new_colnames, all of which obey the conventions described above. Clean up the column names of weather6 by assigning new_colnames to names(weather6)
Replace all empty strings in the events column of weather6 with "None"
One last time, print out the first 6 rows of the weather6 data frame to see the changes
```{r}
new_colnames <- c("date",	"events",
"cloud_cover",	"max_dew_point_f",
"max_gust_speed_mph",	"max_humidity",
"max_sea_level_pressure_in",	"max_temperature_f",
"max_visibility_miles",	"max_wind_speed_mph",
"mean_humidity",	"mean_sea_level_pressure_in",
"mean_temperature_f",	"mean_visibility_miles",
"mean_wind_speed_mph",	"mean_dew_point_f",
"min_dew_point_f",	"min_humidity",
"min_sea_level_pressure_in",	"min_temperature_f",
"min_visibility_miles",	"precipitation_in",
"wind_dir_degrees")

# Clean up column names
names(weather6) <- new_colnames

# Replace empty cells in events column
weather6$events[weather6$events == ""] <- "None"
    
# Print the first 6 rows of weather6
head(weather6)
```

