---
title: "Data Manipulation in R with dplyr"
output: html_notebook
author:  Jeff Gross
---

#Load the dplyr and hflights package

```{r}
options(tibble.print_max = Inf)
options(tibble.width = Inf)

#get a def of a function
getAnywhere(draw_roc_lines)

#install packages
install.packages("dplyr")
install.packages("hflights")
install.packages("data.table")
install.packages("dbplyr")
install.packages("RMySQL")
install.packages("DBI")

# Load the DBI package
library(DBI)

# Load the RMySQL package
library(RMySQL)

# Load the dbplyr package
library(dbplyr)

# Load the data,table package
library(data.table)

# Load the dplyr package
library(dplyr)

# Load the hflights package
library(hflights)

# Call both head() and summary() on hflights
head(hflights)
summary(hflights)
```

#Changing labels of hflights, part 1 of 2
```{r}
# Both the dplyr and hflights packages are loaded into workspace
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Use lut to translate the UniqueCarrier column of hflights
hflights$Carrier <- lut[hflights$UniqueCarrier]

# Inspect the resulting raw values of your variables
glimpse(hflights)
```

#Changing labels of hflights, part 2 of 2
```{r}
# The hflights tbl you built in the previous exercise is available in the workspace.

# Build the lookup table: lut
lut <- c("A"="carrier","B"="weather", "C"="FFA", "D"="security","E"="not cancelled")

# Use the lookup table to create a vector of code labels. Assign the vector to the CancellationCode column of hflights
hflights$Code <- lut[hflights$CancellationCode]


# Inspect the resulting raw values of your variables
glimpse(hflights)
```

#Choosing is not losing! The select verb
```{r}
# hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl with the four columns of hflights related to delay
select(hflights, ActualElapsedTime, AirTime, ArrDelay ,DepDelay)

# Print out the columns Origin up to Cancelled of hflights
select(hflights, Origin:Cancelled)

# Answer to last question: be concise!
select(hflights, 1:4, 12:21)
```
#Helper functions for variable selection

starts_with("X"): every name that starts with "X",
ends_with("X"): every name that ends with "X",
contains("X"): every name that contains "X",
matches("X"): every name that matches "X", where "X" can be a regular expression,
num_range("x", 1:5): the variables named x01, x02, x03, x04 and x05,
one_of(x): every name that appears in x, which should be a character vector.
```{r}
# As usual, hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl containing just ArrDelay and DepDelay
select(hflights, contains("Delay"))

# Print out a tbl as described in the second instruction, using both helper functions and variable names
select(hflights, contains("Num"), contains("Canc"),UniqueCarrier)

# Print out a tbl as described in the third instruction, using only helper functions.
select(hflights, ends_with("Time"), ends_with("Delay"))
```
#Comparison to base R

```{r}
# both hflights and dplyr are available

# Finish select call so that ex1d matches ex1r
ex1r <- hflights[c("TaxiIn", "TaxiOut", "Distance")]
ex1d <- select(hflights, starts_with("Taxi"), Distance)

# Finish select call so that ex2d matches ex2r, "-" removes columns
ex2r <- hflights[c("Year", "Month", "DayOfWeek", "DepTime", "ArrTime")]
ex2d <- select(hflights, Year:ArrTime, -DayofMonth)

# Finish select call so that ex3d matches ex3r
ex3r <- hflights[c("TailNum", "TaxiIn", "TaxiOut")]
ex3d <- select(hflights, starts_with("T"))
```
#Mutating is creating

```{r}
# hflights and dplyr are loaded and ready to serve you.

# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights, ActualGroundTime= ActualElapsedTime - AirTime)

# Add the new variable GroundTime to g1. Save the result as g2.
g2 <- mutate(g1, GroundTime = TaxiIn + TaxiOut)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2, AverageSpeed = Distance / AirTime * 60)

# Print out g3
print(g3)
```
#Logical operators

x < y, TRUE if x is less than y
x <= y, TRUE if x is less than or equal to y
x == y, TRUE if x equals y
x != y, TRUE if x does not equal y
x >= y, TRUE if x is greater than or equal to y
x > y, TRUE if x is greater than y
x %in% c(a, b, c), TRUE if x is in the vector c(a, b, c)
```{r}
# hflights is at your disposal as a tbl, with clean carrier names

# All flights that traveled 3000 miles or more
filter(hflights, Distance >= 3000)

# All flights flown by one of JetBlue, Southwest, or Delta
filter(hflights, UniqueCarrier %in% c("JetBlue", "Southwest", "Delta"))

# All flights where taxiing took longer than flying
filter(hflights, TaxiIn + TaxiOut > AirTime)
```
#Combining tests using boolean operators

R also comes with a set of boolean operators that you can use to combine multiple logical tests into a single test. These include & (and), | (or), and ! (not). Instead of using the & operator, you can also pass several logical tests to filter(), separated by commas. The following two calls are completely equivalent:

filter(df, a > 0 & b > 0)
filter(df, a > 0, b > 0)
```{r}
# hflights is at your service as a tbl!

# All flights that departed before 5am or arrived after 10pm
filter(hflights, DepTime < 500 | ArrTime > 2200)

# All flights that departed late but arrived ahead of schedule
filter(hflights, DepDelay > 0 & ArrDelay < 0)

# All flights that were cancelled after being delayed
filter(hflights, Cancelled == 1 & DepDelay> 0)
```

#Blend together what you've learned!
```{r}
# hflights is already available in the workspace

# Select the flights that had JFK as their destination: c1
c1 <- filter(hflights, Dest == "JFK")

# Combine the Year, Month and DayofMonth variables to create a Date column: c2
c2 <- mutate(c1, Date=paste(Year, Month, DayofMonth, sep = "-"))

# Print out a selection of columns of c2
select(c2, Date, DepTime, ArrTime, TailNum)
```

#Arranging your data
```{r}
# dplyr and the hflights tbl are available

# Definition of dtc
dtc <- filter(hflights, Cancelled == 1, !is.na(DepDelay))

# Arrange dtc by departure delays
arrange(dtc, DepDelay)

# Arrange dtc so that cancellation reasons are grouped
arrange(dtc, CancellationCode)

# Arrange dtc according to carrier and departure delays
arrange(dtc, UniqueCarrier, DepDelay)

```

#Reverse the order of arranging
```{r}
# dplyr and the hflights tbl are available

# Arrange according to carrier and decreasing departure delays
arrange(hflights, UniqueCarrier, desc(DepDelay))

# Arrange flights by total delay (normal order).
arrange(hflights, DepDelay + ArrDelay)

# Keep flights leaving to DFW before 8am and arrange according to decreasing AirTime 
arrange((filter(hflights, Dest == "DFW", DepTime < 800)), desc(AirTime))
```

#The syntax of summarise
```{r}
# hflights and dplyr are loaded in the workspace

# Print out a summary with variables min_dist and max_dist
summarise(hflights, min_dist=min(Distance), max_dist=max(Distance))

# Print out a summary with variable max_div
summarise(filter(hflights, Diverted == 1), max_div=max(Distance))
```

#Aggregate functions

min(x) - minimum value of vector x.
max(x) - maximum value of vector x.
mean(x) - mean value of vector x.
median(x) - median value of vector x.
quantile(x, p) - pth quantile of vector x.
sd(x) - standard deviation of vector x.
var(x) - variance of vector x.
IQR(x) - Inter Quartile Range (IQR) of vector x.
diff(range(x)) - total range of vector x.
```{r}
# hflights is available

# Remove rows that have NA ArrDelay: temp1
temp1 <- filter(hflights, !is.na(ArrDelay))

# Generate summary about ArrDelay column of temp1
summarise(temp1, earliest=min(ArrDelay), average=mean(ArrDelay), latest=max(ArrDelay), sd=sd(ArrDelay))

# Keep rows that have no NA TaxiIn and no NA TaxiOut: temp2
temp2 <- filter(hflights, !is.na(TaxiIn), !is.na(TaxiOut))

# Print the maximum taxiing difference of temp2 with summarise()
summarise(temp2, max_taxi_diff=max(abs(TaxiIn-TaxiOut)))
```

#dplyr aggregate functions

dplyr provides several helpful aggregate functions of its own, in addition to the ones that are already defined in R. These include:

first(x) - The first element of vector x.
last(x) - The last element of vector x.
nth(x, n) - The nth element of vector x.
n() - The number of rows in the data.frame or group of observations that summarise() describes.
n_distinct(x) - The number of unique values in vector x.
```{r}
# hflights is available with full names for the carriers

# Generate summarizing statistics for hflights
summarise(hflights,
          n_obs = n(),
          n_carrier = n_distinct(UniqueCarrier),
          n_dest = n_distinct(Dest))

# All American Airline flights
aa <- filter(hflights, UniqueCarrier == "American")

# Generate summarizing statistics for aa 
summarise(aa, n_flights=n(),
                n_canc = sum(Cancelled == 1),
                avg_delay = mean(ArrDelay, na.rm=TRUE))
```

#Overview of syntax
Take the hflights data set and then ...
Add a variable named diff that is the result of subtracting TaxiIn from TaxiOut, and then ...
Pick all of the rows whose diff value does not equal NA, and then ...
Summarise the data set with a value named avg that is the mean diff value.
```{r}
# hflights and dplyr are both loaded and ready to serve you

# Write the 'piped' version of the English sentences.
hflights %>%
  mutate(diff=TaxiOut - TaxiIn) %>%
  filter(!is.na(diff)) %>%
  summarise(avg = mean(diff))
```
#Drive or fly? Part 1 of 2
mutate() the hflights dataset and add two variables:
     RealTime: the actual elapsed time plus 100 minutes (for the overhead that flying involves) and
     mph: calculated as Distance / RealTime * 60, then
filter() to keep observations that have an mph that is not NA and that is below 70, finally
summarise() the result by creating four summary variables:
     n_less, the number of observations,
     n_dest, the number of destinations,
     min_dist, the minimum distance and
     max_dist, the maximum distance.
```{r}
# Chain together mutate(), filter() and summarise()
hflights %>%
  mutate(RealTime = ActualElapsedTime + 100) %>%
  mutate(mph = Distance / RealTime * 60) %>%
  filter(!is.na(mph) & mph < 70) %>%
  summarise(n_less = n(),
            n_dest = n_distinct(Dest),
            min_dist = min(Distance),
            max_dist = max(Distance))
  
```

#Drive or fly? Part 2 of 2
filter() the result of mutate to:
keep observations that have an mph under 105 or for which Cancelled equals 1 or for which Diverted equals 1.
summarise() the result by creating four summary variables:
n_non, the number of observations,
n_dest, the number of destinations,
min_dist, the minimum distance and
max_dist, the maximum distance.
```{r}
# Finish the command with a filter() and summarise() call
hflights %>%
  mutate(RealTime = ActualElapsedTime + 100, mph = Distance / RealTime * 60) %>%
  filter(mph < 105 | Cancelled == 1 | Diverted == 1) %>%
  summarise(n_non = n(),
            n_dest = n_distinct(Dest),
            min_dist = min(Distance),
            max_dist = max(Distance))
```

#Advanced piping exercise
filter() the hflights tbl to keep only observations whose DepTime is not NA, whose ArrTime is not NA and for which DepTime exceeds ArrTime.
Pipe the result into a summarise() call to create a single summary variable: num, that simply counts the number of observations
```{r}
# hflights and dplyr are loaded

# Count the number of overnight flights
hflights %>%
  filter(!is.na(DepTime) & !is.na(ArrTime) & DepTime > ArrTime) %>%
  summarise(num=n())
```
#Unite and conquer using group_by
Use group_by() to group hflights by UniqueCarrier.
summarise() the grouped tbl with two summary variables:
p_canc, the percentage of cancelled flights
avg_delay, the average arrival delay of flights whose delay does not equal NA.
Finally, order the carriers in the summary from low to high by their average arrival delay. Use percentage of flights cancelled to break any ties.
```{r}
# hflights is in the workspace as a tbl, with translated carrier names

# Make an ordered per-carrier summary of hflights
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise(p_canc = mean(Cancelled == 1)*100,
            avg_delay = mean(ArrDelay, na.rm=TRUE)) %>%
  arrange(avg_delay, p_canc)
```
#Combine group_by with mutate
filter() the hflights tbl to only keep observations for which ArrDelay is not NA and positive.
Use group_by() on the result to group by UniqueCarrier.
Next, use summarise() to calculate the average ArrDelay per carrier. Call this summary variable avg.
Feed the result into a mutate() call: create a new variable, rank, calculated as rank(avg).
Finally, arrange by this new rank variable
```{r}
# dplyr is loaded, hflights is loaded with translated carrier names

# Ordered overview of average arrival delays per carrier
hflights %>%
  filter(!is.na(ArrDelay), ArrDelay>0) %>%
  group_by(UniqueCarrier) %>%
  summarise(avg = mean(ArrDelay)) %>%
  mutate( rank = rank(avg)) %>%
  arrange(rank)
```
#Advanced group_by exercises
How many airplanes flew to only one destination? The tbl you print out should have a single column, named nplanes and a single row.
Find the most visited destination for each carrier. The tbl you print out should contain four columns:
UniqueCarrier and Dest,
n, how often a carrier visited a particular destination,
rank, how each destination ranks per carrier. rank should be 1 for every row, as you want to find the most visited destination for each carrier.
```{r}
# dplyr and hflights (with translated carrier names) are pre-loaded

# How many airplanes only flew to one destination?
hflights %>%
  group_by(TailNum) %>%
  summarise(ndest = n_distinct(Dest)) %>%
  filter(ndest == 1) %>%
  summarise(nplanes=n())

# Find the most visited destination for each carrier
hflights %>%
  group_by(UniqueCarrier, Dest) %>%
  summarise(n=n()) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
```

#dplyr deals with different types
Use summarise() to calculate n_carrier, the total number of unique carriers in hflights2. Whether or not you use the pipe is up to you!
```{r}
library(data.table)
hflights2 <- as.data.table(hflights)

# hflights2 is pre-loaded as a data.table

# Use summarise to calculate n_carrier
hflights2 %>%
  summarise(n_carrier = n_distinct(UniqueCarrier))
```
#dplyr and mySQL databases
Try to understand the code that creates nycflights, a reference to a MySQL table.
Use glimpse() to check out nycflights. Although nycflights is a reference to a tbl in a remote database, there is no difference in syntax. Look carefully: the variable names in nycflights differ from the ones in hflights!
Group nycflights data by carrier, then summarise() with two variables: n_flights, the number of flights flown by each carrier and avg_delay, the average arrival delay of flights flown by each carrier. Finally, arrange the carriers by average delay from low to high.
```{r}
# Set up a connection to the mysql database
my_db <- src_mysql(dbname = "dplyr", 
                   host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                   port = 3306, 
                   user = "student",
                   password = "datacamp")

# Reference a table within that source: nycflights
nycflights <- tbl(my_db, "dplyr")

# glimpse at nycflights
glimpse(nycflights)

# Ordered, grouped summary of nycflights
nycflights %>%
  group_by(carrier) %>%
  summarise(n_flights=n(), avg_delay=mean(arr_delay)) %>%
  arrange(avg_delay)
```

