---
title: "Data Visualization with ggplot2"
output: html_notebook
author: Jeff Gross
---
#Libraries, Misc
```{r}
install.packages("ggplot2")

library(ggplot2)

#get a def of a function
getAnywhere(draw_roc_lines)
```


#Exploring ggplot2, part 1

Instructions
Load the ggplot2 package using library(). It is already installed on DataCamp's servers.
Use str() to explore the structure of the mtcars dataset.
Hit Submit Answer. This will execute the example code on the right. See if you can understand what ggplot does with the data.
```{r}
# Load the ggplot2 package
library(ggplot2)

# Explore the mtcars data frame with str()
str(mtcars)

# Execute the following command
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()
```
#Exploring ggplot2, part 2

Instructions
Change the ggplot() command by wrapping factor() around cyl.
Hit Submit Answer and see if the resulting plot is better this time.
```{r}
# Load the ggplot2 package
library(ggplot2)

# Change the command below so that cyl is treated as factor
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_point()
```
#Exploring ggplot2, part 3

Instructions
ggplot2 has already been loaded for you. Take a look at the first command. It plots the mpg (miles per galon) against the weight (in thousands of pounds). You don't have to change anything about this command.
In the second call of ggplot() change the color argument in aes() (which stands for aesthetics). The color should be dependent on the displacement of the car engine, found in disp.
In the third call of ggplot() change the size argument in aes() (which stands for aesthetics). The size should be dependent on the displacement of the car engine, found in disp.
```{r}
# A scatter plot has been made for you
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) +
  geom_point()

# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()
```
#Exploring ggplot2, part 4

Instructions
Explore the diamonds data frame with the str() function.
Use the + operator to add geom_point() to the first ggplot() command. This will tell ggplot2 to draw points on the plot.
Use the + operator to add geom_point() and geom_smooth(). These just stack on each other! geom_smooth() will draw a smoothed line over the points.
```{r}
# Explore the diamonds data frame with str()
str(diamonds)

# Add geom_point() with +
ggplot(diamonds, aes(x = carat, y = price))+geom_point()


# Add geom_point() and geom_smooth() with +
ggplot(diamonds, aes(x = carat, y = price))+geom_point()+geom_smooth()
```
#Exploring ggplot2, part 5

Instructions
Plot 2 - Copy and paste plot 1, but show only the smooth line, no points.
Plot 3 - Show only the smooth line, but color according to clarity by placing the argument color = clarity in the aes() function of your ggplot() call.
Plot 4 - Draw translucent colored points.
Copy the ggplot() command from plot 3 (with clarity mapped to color).
Remove the smooth layer.
Add the points layer back in.
Set alpha = 0.4 inside geom_point(). This will make the points 40% transparent.
```{r}
# 1 - The plot you created in the previous exercise
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()

# 2 - Copy the above command but show only the smooth line
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_smooth()
  

# 3 - Copy the above command and assign the correct value to col in aes()
ggplot(diamonds, aes(x = carat, y = price, color=clarity)) +
  geom_smooth()


# 4 - Keep the color settings from previous command. Plot only the points with argument alpha.
ggplot(diamonds, aes(x = carat, y = price, color=clarity)) +
  geom_point(alpha=.4)
```
#Understanding the grammar, part 1

Instructions
Define the data (diamonds) and aesthetics layers. Map carat on the x axis and price on the y axis. Assign it to an object: dia_plot.
Using +, add a geom_point() layer (with no arguments), to the dia_plot object. This can be in a single or multiple lines.
Note that you can also call aes() within the geom_point() function. Map clarity to the color argument in this way.
```{r}
# Create the object containing the data and aes layers: dia_plot
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# Add a geom layer with + and geom_point()
dia_plot + geom_point()

# Add the same geom layer, but with aes() inside
dia_plot + geom_point(aes(color = clarity))
```
#Understanding the grammar, part 2

Instructions
1 - The dia_plot object has been created for you.
2 - Update dia_plot so that it contains all the functions to make a scatter plot by using geom_point() for the geom layer. Set alpha = 0.2.
3 - Using +, plot the dia_plot object with a geom_smooth() layer on top. You don't want any error shading, which can be achieved by setting the se = FALSE in geom_smooth().
4 - Modify the geom_smooth() function from the previous instruction so that it contains aes() and map clarity to the col argument.
```{r}
# 1 - The dia_plot object has been created for you
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# 2 - Expand dia_plot by adding geom_point() with alpha set to 0.2
dia_plot <- dia_plot + geom_point(alpha=.2)

# 3 - Plot dia_plot with additional geom_smooth() with se set to FALSE
dia_plot + geom_smooth(se = FALSE)

# 4 - Copy the command from above and add aes() with the correct mapping to geom_smooth()
dia_plot + geom_smooth(aes(col = clarity), se = FALSE)
```

