---
title: "String Manipulation in R with stringr"
author: Jeff Gross
output: html_notebook
---

#Quotes

####Background: There are cases where you need both ' and " inside the string. In this case, fall back to the first guideline and use " to define the string, but you'll have to escape any double quotes inside the string using a backslash (i.e. \").

Line 1: The table was a large one, but the three were all crowded together at one corner of it:
Line 2: "No room! No room!" they cried out when they saw Alice coming.
Line 3:"There's plenty of room!" said Alice indignantly, and she sat down in a large arm-chair at one end of the table.

```{r}
# Define line1, no quotes inside the text, so just use quotes
line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"

# Define line2, double quotes inside the text, so use single quotes
line2 <- '"No room! No room!" they cried out when they saw Alice coming.'

# Define line3, double & single quotes inside the text, so use escape (\")
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."
```
#What you see isn't always what you have

####Background: You can pass writeLines() a vector of strings and it will print them to the screen, each on a new line. This is a great way to check the string you entered really does represent the string you wanted. The function cat() is very similar to writeLines(), but by default separates elements with a space, and will attempt to convert non-character objects to a string. 

```{r}
install.packages("emojifont")
library(emojifont)

line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"
line2 <- "\"No room! No room!\" they cried out when they saw Alice coming."
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."

# Putting lines in a vector
lines <- c(line1, line2, line3)

# Print lines
lines

# Use writeLines() on lines
writeLines(lines)

# Write lines with a space separator
writeLines(lines, sep=" ")

# Use writeLines() on the string "hello\n\U1F30D"
writeLines("hello\n\U1F30D")

cat("\xa3")
```
#Escape sequences

####Background: \U followed by up to 8 hex digits sequence denotes a particular Unicode character

\n	newline
\r	carriage return
\t	tab
\b	backspace
\a	alert (bell)
\f	form feed
\v	vertical tab
\\	backslash \
\'	ASCII apostrophe '
\"	ASCII quotation mark "
\`	ASCII grave accent (backtick) `
\nnn	character with given octal code (1, 2 or 3 digits)
\xnn	character with given hex code (1 or 2 hex digits)
\unnnn	Unicode character with given code (1--4 hex digits)
\Unnnnnnnn	Unicode character with given code (1--8 hex digits)

Alternative forms for the last two are \u{nnnn} and \U{nnnnnnnn}. All except the Unicode escape sequences are also supported when reading character strings by scan and read.table if allowEscapes = TRUE. Unicode escapes can be used to enter Unicode characters not in the current locale's charset (when the string will be stored internally in UTF-8).

The parser does not allow the use of both octal/hex and Unicode escapes in a single string.

These forms will also be used by print.default when outputting non-printable characters (including backslash).

Embedded nuls are not allowed in character strings, so using escapes (such as \0) for a nul will result in the string being truncated at that point (usually with a warning).

```{r}
# Should display: To have a \ you need \\
writeLines("To have a \\ you need \\\\")

# Should display: 
# This is a really 
# really really 
# long string
writeLines("This is a really\nreally really\nlong string")

# Use writeLines() with 
# "\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e"
writeLines("\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e")
cat("\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e")
```
#Turning numbers into strings

#Using format() with numbers

```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1), digits=1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1), digits=1)

# Format percent_change to one place after the decimal point
format(percent_change, digits=2)

# Format income to whole numbers
format(income, digits=2)
```
#Using format() with numbers

```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1), digits=1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1), digits=1)

# Format percent_change to one place after the decimal point
format(percent_change, digits=2)

# Format income to whole numbers
format(income, digits=2)
```
#Controlling other aspects of the string



